---
title: <center>Clusters on Glom and Tubular Patients</center>
author: <center>Yang Li</center>
date: <center>`r format(Sys.time(), '%m/%d/%Y')`</center>
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    number_sections: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Goal

The primary goal is to generate a heatmap based on the gene expression count matrix and metadata (sample information) for glomerular and tubular patients, respectively. In each heatmap, patient clusters are also presented. 

# Methods

* Package DEseq2 is used to calculate sample-sample distances.
* Package pheatmap is used to generate the heatmap of gene expressions. 
* Package pheatmap is used to perform clustering on patients based on the sample-sample distance matrix. 

# Glomerular Results

## Data Description

* Number of genes: 608
* Number of samples: 89
* Number of sample clusters: 4 (set manually)

## Data Preprocessing

* The sheet "Glom" in the Excel file is merged as an expression matrix (608 x 89).

## Heatmap

```{r,echo=F,eval=T,message=F,warning=F,error=F}
setwd("/Users/liyang1/Box/OSU/DR-ALAMAANI/")
Count.Mat <- read.csv("Glom_Count.csv", header = T, row.names = 1) # Read read count table in .csv file
# into program
Gene.Names <- rownames(Count.Mat)
Gene.Names.Uniq <- unique(Gene.Names)

# Check whether there exist duplicated gene names
if (length(Gene.Names) != length(Gene.Names.Uniq))
{
  stop("Error: There exist duplicated gene names.")
}

# Check whether there exist duplicated sample names
Sample.Names <- colnames(Count.Mat)
Sample.Names.Uniq <- unique(Sample.Names)

if (length(Sample.Names) != length(Sample.Names.Uniq))
{
  stop("Error: There exist duplicated sample names.")
}

Meta.Mat <- read.csv("Glom_Metadata.csv", header = T,row.names = 1) # Read meta data in .csv file

if(length(Sample.Names) != length(rownames(Meta.Mat)))
{
  stop("Error: The number of samples in count matrix file and meta file is not consistent.")
}

library(DESeq2)

dds.Obj <- DESeqDataSetFromMatrix(countData = Count.Mat, colData = Meta.Mat, design=~Time, tidy=F)
# Transform the count matrix and metadata into dds object
dds.Obj <- dds.Obj[rowSums(counts(dds.Obj))>1,]
# Filter the data
vsd <- varianceStabilizingTransformation(dds.Obj, blind = F) # Standardize the data
Sample.Dist <- dist(t(assay(vsd))) # Calculate the distance between groups
# hc <- hclust(Sample.Dist, method = "ward.D2") # Hierarchical clustering 
library(pheatmap)
list = pheatmap(mat = Sample.Dist, cluster_rows = T, cluster_cols = T, 
         annotation_col = Meta.Mat)
row_cluster = cutree(list$tree_row,k=4)

```

## Clustering Results

```{r,echo=F,eval=T,message=F,warning=F,error=F}
Cluster.Table <- data.frame(row_cluster, Meta.Mat[["Name"]])
library("DT")
datatable(Cluster.Table, colnames = c("ID", "Cluster", "Name"), 
          extensions = c('FixedColumns','Buttons'), 
          options = list(
            pageLength = 15,
            scrollX = TRUE,
            scrollCollapse = TRUE,
            dom = 'Bfrtip',
            buttons = c('copy', 'csv', 'excel')
          ))
```

# Tubular Results

## Data Description

* Number of genes: 598
* Number of samples: 113
Number of clusters: 4 (set manually)

## Data Preprocessing

* The sheet "Tubular" in the Excel file is merged as an expression matrix (598 x 113).
* The first three samples "36AT", "1AT" and "38AT" are ruled out from analysis, since many genes are not expressed in the three samples.

## Heatmap

```{r,echo=F,eval=T,message=F,warning=F,error=F}
setwd("/Users/liyang1/Box/OSU/DR-ALAMAANI/")
Count.Mat <- read.csv("Tubular_No_Triple_Count.csv", header = T, row.names = 1) 
# Read read count table in .csv file
# into program
Gene.Names <- rownames(Count.Mat)
Gene.Names.Uniq <- unique(Gene.Names)

# Check whether there exist duplicated gene names
if (length(Gene.Names) != length(Gene.Names.Uniq))
{
  stop("Error: There exist duplicated gene names.")
}

# Check whether there exist duplicated sample names
Sample.Names <- colnames(Count.Mat)
Sample.Names.Uniq <- unique(Sample.Names)

if (length(Sample.Names) != length(Sample.Names.Uniq))
{
  stop("Error: There exist duplicated sample names.")
}

Meta.Mat <- read.csv("Tubular_No_Triple_Metadata.csv", header = T,row.names = 1) 
# Read meta data in .csv file

if(length(Sample.Names) != length(rownames(Meta.Mat)))
{
  stop("Error: The number of samples in count matrix file and meta file is not consistent.")
}

library(DESeq2)

dds.Obj <- DESeqDataSetFromMatrix(countData = Count.Mat, colData = Meta.Mat, design=~Time, tidy=F)
# Transform the count matrix and metadata into dds object
dds.Obj <- dds.Obj[rowSums(counts(dds.Obj))>1,]
# Filter the data
vsd <- varianceStabilizingTransformation(dds.Obj, blind = F) # Standardize the data
Sample.Dist <- dist(t(assay(vsd))) # Calculate the distance between groups
# hc <- hclust(Sample.Dist, method = "ward.D2") # Hierarchical clustering 
library(pheatmap)
list = pheatmap(mat = Sample.Dist, cluster_rows = T, cluster_cols = T, 
         annotation_col = Meta.Mat)
row_cluster = cutree(list$tree_row,k=4)

```

## Clustering Results

```{r,echo=F,eval=T,message=F,warning=F,error=F}
Cluster.Table <- data.frame(row_cluster, Meta.Mat[["Name"]])
library("DT")
datatable(Cluster.Table, colnames = c("ID", "Cluster", "Name"), 
          extensions = c('FixedColumns','Buttons'), 
          options = list(
            pageLength = 15,
            scrollX = TRUE,
            scrollCollapse = TRUE,
            dom = 'Bfrtip',
            buttons = c('copy', 'csv', 'excel')
          ))
```